Response to reviewers

### 2024-11-28 Manuscript/ecography_reponse_letter.tex
Update ecography_reponse_letter.tex

% \textbf{Reviewer: 2}

% Upon reading the title and introduction, I was intrigued and had high expectations for the proposed ``what, how, why'' framework of trait-based analysis
% However, after thoroughly reviewing the manuscript, it became clear that the study does not fully deliver on its initial promise of providing a guideline for applying this framework to ecological data analysis
% Instead, it focuses more on evaluating the varying responses of different functional diversity indices when specific data points are selectively removed from a dataset
% The primary outcome is that different functional diversity indices exhibit distinct and often unpredictable responses to data removal
% However, this observation, while valid, does not seem to offer significant new insights into ecological processes
% In essence, the study highlights the dependence of analytical outcomes on the choice of index---a finding that, while true, feels somewhat self-evident.

% It is possible that these concerns arise due to insufficient descriptions of the methodology and inconsistencies in terminology (specific points are detailed below)
% As the manuscript stands, I am unable to recommend it for publication
% That said, I have outlined several constructive suggestions that I hope will assist the authors in revising and refining the paper.

% \textbf{General comments}

% \textbf{Main model idea}
%  The paper falls short in demonstrating how the ``what, how, why'' framework can be applied to ecological data
% This issue may stem from examining two distinct layers: the simple removal of points from datasets and the subsequent interpretation of these removals in terms of ecological processes, which are not explicitly simulated in the model
% The distinction between these layers is confusing and is primarily addressed in Table 1, which does not effectively clarify these points
% To improve clarity, I recommend a more systematic presentation of these issues in the main text: This would involve, describing the relevant ecological processes (e.g., facilitation, filtering, competition), explaining how these processes influence species-associated trait distributions, and then outlining how these influences are represented in the five trait-removal processes.

@@

% \textbf{Model description:}
%  The section on simulating trait-space patterns needs to be significantly expanded to include more technical details regarding how the simulations were conducted
% Key information, such as the model parameters, is entirely absent
% A more thorough explanation of the model setup, parameters, and procedures is required to ensure that the simulations can be reproduced without needing to consult external publications (such as those by Guillerme and Mammola)
% To address this, I suggest adding a detailed explanation of the model in either the main Methods section or in a supplementary appendix, to ensure that all necessary parameters and computational steps are clearly specified.

@@

% \textbf{Consistent terminology:}
%  The paper's clarity is hindered by inconsistent, imprecise, and shifting terminology, making it unnecessarily difficult to follow
% For example, the entities generated by the model are referred to interchangeably as ``elements,'' ``traits,'' ``tips,'' or ``species,'' with no clear rationale for the variation
% Similar, for many other terms that are often used inconsistently or fuzzy (see detailed comments below)
% This lack of consistency creates confusion and often forced me to repeatedly revisit sections of the text to discern the intended meaning
% To improve readability, I recommend establishing a consistent terminology early in the paper and adhering to it throughout
% This would significantly reduce ambiguity and enhance the reader's ability to follow the methodology and conclusions.
%  We have now standardised the terminology following \textbackslash cite\{guillerme2020shifting\} and \textbackslash cite\{mammola2021concepts\}
% In brief, elements, tips or species are now referred to as ``observations''; and traits, dimensions and columns as ``dimensions'' when referring to our simulation approach
% We also added a Glossary to the supplementary materials @@@ for clarity.

% \textbf{Detailed comments:}

% L114: Really ``A single Brownian motion trait''? This implies that only a single trait dimension is generated, whereas from the later text it seems that higher-dimensional traits are simulated.
%  By ``trait'' here we refer to the single process (Brownian Motion) which can be simulated in any number of dimensions (i.e.~generating a time dependent element value using 2, 4 or 8 dimension coordinates to describe it)
% In this context we differentiate ``dimensions'' which can be interpreted as a column in the data matrix, and the ``trait'' which is the definition of the measured (or simulated) feature
% For example, the trait ``geographical location'' can be described using two dimensions (latitude and longitude) or can be described as two traits (``latitude coordinate'' and ``longitude coordinate'')
% In this manuscript we preferred using the first definition
% We've clarified this as follows:
%  \textit{First, we simulated a single random time dependent trait (Brownian Motion) under a model where lineages only speciate (no extinction; i.e.~a pure birth speciation model) until reaching 200 elements in \textbackslash texttt\{R\} \textbackslash citep\{rcore\} using \textbackslash texttt\{treats\} \textbackslash citep\{guillerme2024treats\}
% This trait was simulated either as a 2, 4 or 8 dimensional trait, in other words a Brownian Motion simulated in a 2, 4 or 8 dimensional space.}
%  Lines@@@

% L115: what is a ``tip''?
%  By tip we meant a node with no descendants in an evolutionary tree
% We have now standardised the terminology (see response to general comment above).

% L115, ``200 elements'': Why are so few data points used? A dataset of only 200 elements (reduced to 40 after data removal) seems exceptionally sparse, especially in a 4- or 8-dimensional space, as noted later in the text
% What is the motivation behind studying such a sparse space? One of the main advantages of using simple models like a branching process is that they allow for the generation of larger datasets, which would provide more robust statistics and better analytical power.
%  @@@TODO: Rerun the simulations with 400, 800 and 1000 elements.
%  But also highlight that 40 species is already quite big
% We can also link that to the use of FD in palaeontology (disparity) where 40 is also already quite big!

% LL 120-124, ``four different stressors\ldots{} five specific algorithms'': Inconsistency in the used terminology
% In this section, ``stressors'' appear to refer to the amount of data being removed (ranging from 20\% to 80\%), but later, the term ``stressors'' is used to describe the five data-removal processes---referred to as ``specific algorithms'' here.
%  We have now standardised the terminology throughout the manuscript (and included it in the glossary table in the supplementary materials): ``algorithm'' now designate the code algorithm, ``stressors'' (and ``stressors' intensity'') now designates how we modify the data (using the algorithms mentioned above) and ``mechanisms'' now designates the biological mechanisms we are approximating using a specific stressor (that is based on a specific algorithm)
% We've namely changed this section for clarity:
%  \textit{We then applied five different stressors to the trait space with different intensities.}
%  \textit{The five different stressors are effectively five specific algorithms removing either 20\textbackslash\%, 40\textbackslash\%, 60\textbackslash\% or 80\textbackslash\% of the data (the intensities of the stressors, resulting in trait spaces with, respectively, 160, 120, 80 and 40 observations).}
%  \textit{We used the following specific algorithms (Figure \textbackslash ref\{Fig:simulations\}; Table \textbackslash ref\{Tab:mechanisms\}; all algorithms, except ``increasing evenness'\,', were previously described in \textbackslash cite\{guillerme2020shifting\}):}
%  Lines @@@

% L130, ``increasing density'': What is the motivation for this term? While I understand that this removal scheme may result in an overall increase in average density, the removal itself is not directly based on density but rather on pairwise distance in trait space
% For instance, the equalizing scheme would also lead to an increase in density.
%  We used the term ``increasing density'' here to differentiate from ``decreasing size''
% The reviewer is totally correct in their understanding that, especially for normally distributed data, changes in density in the data actually occur to some extent with all the algorithms used here
% The stressor names and their algorithms are based on Guillerme et al 2020 and their implementation in the dispRity package: in their description they don't assume any specific distribution of the data a priori
% In fact, when using a uniform distribution, the decreasing size stressor does not change the density and the increasing density algorithm does not decrease the size of the trait space.
%  We've now added the following note to include the reviewer's comment:
%  \textit{Note that the algorithm used here to simulate the stressors depend, to some extent, to the distribution of the data for each trait space.}
%  \textit{The algorithms used can lead to similar stressors effects depending on the simulated data distribution.}
%  \textit{For example, ``increasing density'\,' and ``decreasing size'\,' stressors will lead to just what their name suggest if the observations are uniformly distributed (just average increase in density for ``increasing density'\,' and just overall decrease in size for the ``decreasing size'\,' stressors) but will have more or less similar effects (albeit no identical) in the case of a normal distribution (average increase in density and overall decrease in size for both ``increasing density'\,' and ``decreasing size'\,' stressors.)}
%  Lines @@@

% L134, ``shifting space'': Similar to the previous point, this term is somewhat misleading
% This removal scheme does not explicitly shift the distribution; rather, it cuts the distribution at one edge, causing the mean to shift
% However, this effect of shifting the mean is not unique to this scheme---other removal schemes will generally result in a shift of the mean as well.
%  This is also a fair point
% See comment above.

% L135, ``with the maximal value'': What does ``maximal value'' mean for points in a higher-dimensional space? Maybe this refer to the location in trait space with the highest density of trait elements?
%  We have now specified what we meant by maximal value:
%  \textit{Shifting space: by removing the required amount of data from a distance (radius) \(
% rho'\) of the observation with the maximum numerical value on all dimensions.}
%  \textit{This is similar to the decreasing size algorithm but instead of choosing \(
% rho\) to be the radius from the centre of the trait space, \(
% rho'\) is a radius from the ``top right corner'\,' of the trait space.}
%  \textit{In other words the centre of the radius \(
% rho'\) is the observation with the highest numerical value on all dimensions - in a 2D representation this is the observation in the top right corner (with \(
% rho'\_{20} \> 
% rho'\_{40} \> 
% rho'\_{60} \> 
% rho'\_{80}\)).}
%  Lines @@@

% LL138- 145: I found this section on the competition method difficult to follow:
%  - The term ``bandwidth'' is unclear
% What does ``bandwidth'' refer to in this context? Does it imply that resampling was done using a frequency-based method, or is it referring to a specific range or interval of trait values in the distribution? This ambiguity ties back to my earlier request for more technical details regarding the simulation process.
%  - I am also unclear on the meaning of ``n''
% Could you clarify what ``n'' represents here? For example, what is meant by the ``probability of sampling of n'' and the reference to ``high/low values of n''? Without this information, it is difficult to grasp how the competition process is functioning.
%  - The term ``flattening the curve'' needs more context
% What curve is being flattened, and how does this relate to the competition method?
%  We have now completely rewritten the description of this stressor by defining more clearly what the bandwidth is, renaming the variables to be easier to understand, removed mention of flattening the curve and rewrote the equations to be specific to the increase in evenness (previous equations were generalised to both increase or decrease the eveness depending on the variable \textit{i} (formerly \textit{n}) being now always a proportion of removal (e.g.~0.2, 0.4, \ldots{} - former description worked with \textit{n} being also a proportion to add e.g.~1.2, 1.4, etc\ldots):
%  \textit{Increasing evenness: by resampling the proportion of data (i.e.~20\textbackslash\%, 40\textbackslash\%, 60\textbackslash\% or 80\textbackslash\%) but with skewed resampling probabilities.}
%  \textit{In brief, this algorithm reduces the probability of resampling observations in regions of the trait space that have many observations and increases it in regions that have few observations.}
%  \textit{This is done by selecting of \(B\) discrete categories to summarise the distribution - i.e.~bandwidths - using Silverman's ``rule of thumb'\,' \textbackslash citep{[}\textbackslash texttt\{bw.nrd0()\} function in \textbackslash texttt\{R\};{]}{[}{]}\{silverman1986density\}.}
%  \textit{Each of these discrete category \(B\) has an observed probability of sampling of \(b\) (the proportion of observations in the category) and each observation then gets a probability of resampling scaling with the intensity of the stressor (the proportion of what to remove \(i\)) of \(i 
% times (1-b)^{p}\).}
%  \textit{Where \(p\) is factor increasing the scaling power of the algorithm (here we used \(p=3\)).}
%  \textit{In other words, observations in categories with a low density (\(b\<0.5\)) were more likely to be resampled and the ones in categories with high density (\(b\>0.5\)) were less likely to be resampled.}
%  \textit{This approximates our \textbackslash textbf\{competition mechanism\}, were observations in dense regions of the trait space are more likely to be removed than in sparse regions of the trait space.}
%  Lines @@@

% Figure 1: The caption is confusing and needs significant improvement
% Below are some specific points that require attention:
%  - The caption completely omits any description of what is shown in left and right columns (i.e., 1d and 2d case).
%  We have now clarified what each column contains in the figure caption.
%  - It is unclear whether this figure (both the 1d and 2d cases) is based on the actual simulations or is merely a schematic sketch
% For instance, in the right column, it seems unlikely that the figure actually depicts 200 points.
%  The figures do depict 200 points (200 blue in row A and 100 blue + 100 orange in the other columns)
% We have now specified this in the caption.
%  - The caption mentions a ``trait space of 2, 4, or 8 dimensions,'' yet a 1d trait space is shown
% If a 1d trait space is an illustrative example, this should be specified in the caption.
%  We have now specified that we are showing 1 or 2 dimensions regardless of the number of dimensions per traits (i.e.~8 dimensional traits are not fully represented here but what's happening on their first and second dimensions is displayed here)
% We think one confusion on our part is the distinction between a dimension (a column in a matrix) and a trait (a measurable feature for an observation)
% In our case we are always only looking at one trait per trait space
% Although these traits can have 2, 4 or 8 dimensions
% We have now clarified this in the glossary.
%  - The pattern shown in the subfigure labeled ``2d facilitating mechanism'' appears questionable
% If point-pairs from a 2d distribution are removed based on a threshold distance, the resulting pattern would not typically resemble what is depicted here.
%  The illustration of the 2D facilitating mechanism directly comes from the output of the dispRity::reduce.space function where the algorithm is implemented
% Here it selected every \textit{n} pairs of observations that are the closest to each other so that 50\% of the data is selected
% Here \textit{n} is 50 so that 100 observations are selected
% The algorithm is recursive so that once a pair of observations is selected, it cannot be selected again.
%  - Terminology: The use of the term ``species'' in the caption is confusing, as the methods seem to generate traits rather than species.
%  We have now removed any mention of species.
%  - The phrase ``removing half of the species, resulting in two groups of tips'' is unclear
% How are these two groups defined? Presumably, one group consists of the removed traits and the other of non-removed traits
% Additionally, it is unclear which color (blue or orange) corresponds to which group.
%  We have now specified this in the new caption.
%  - I am skeptical about the use of ecological terminology such as ``competition'' and ``facilitation'' in this context
% What is actually being shown are data distributions resulting from different point-removal algorithms, not ecological interactions per se
% While these processes can be discussed and associated with ecological mechanisms in the text or discussion (if well justified), using such terms directly in the figure caption is highly misleading.
%  @@@TODO: Links back to main comment from reviewer 2: Go back to the previous version of table 1 that gives more description of the models
% One thing here is probably to justify more why we think that, for example, ``removing points from the side of a traitspace = filtering''.
%  We now changed the labels in the figure and the caption to highlight that we are indeed approximating such mechanisms, not simulating them per se as the reviewers pointed out.

% LL 150-152: The description is unclear and appears to use the term ``simulation'' inconsistently
% In the phrase ``replicates per simulations,'' the term ``simulation'' seems to have a different meaning than in ``smallest number of simulations.'' Additionally, the first part of the sentence suggests that variance decreases with an increasing number of simulations, whereas the second part implies that adding more simulations leads to increased variance.
%  There was a confusion indeed
% We now refer to ``the simulation pipeline'' for the whole set of replicates per stressors, and ``replicates'' for each individual run (one trait space + one stressor)
% We changed this sentence to:
%  \textit{We used 171 replicates because that was empirically the smallest number of replicates required to reach a variance between replicates lower than 1\textbackslash\% across all metrics (i.e.~any additional replicates beyond 171 added less that 1\textbackslash\% extra variance).}
%  Lines @@@

% L 166: what does ``uncorrelated'' mean here? Does this mean that the branching process was independently repeated four times, once for each trait dimension, or does it mean that it generates four trait dimensions simultaneously, with these dimensions being uncorrelated?
%  We meant that the dimensions for each multidimensional traits are uncorrelated (i.e.~the trait is a uncorrelated Brownian motion process in n dimensions)
% We have now clarified this point:
%  \textit{We repeated each simulation pipeline for 4 dimensional traits (generate a trait space and apply the stressor - Figure \textbackslash ref\{Fig:simulation\_results\}) for 2 and 8 dimensional traits with not correlation between each dimensions in each simulation pipeline (supplementary materials).}
%  Lines @@@

% LL 170-178: why then did the authors not simply generate more data points (see my comment above)?
%  @@@TODO: Rerun the simulations with 400, 800 and 1000 elements.

% Table 1: The integration of this table into the main text could be improved
% Specifically, the abrupt mention of species (which are not part of the model) is confusing.
%  We have now improved the context of the tables with predictions and removed most of the mentions of ``species'' now replaced by the standardised term of ``observations'' (unless they were specific examples).

% L 178 and several instances later ``o'' should read ``0'' (zero)
%  These integers were compiled using the latex mathpazo font
% We have now clarified them by changing into a default serif font.

% L184, I suggest to mention here, that these functional metrics are denoted as trait space occupancy later on.
%  Great suggestions
% We added the following:
%  \textit{We structured our simulations based on three aspects of diversity commonly captured by functional diversity metrics (trait space occupancy metrics; \textbackslash cite\{mammola2021concepts\}) using 12 metrics (Tab
% \textbackslash ref\{Tab:metrics\}):}
%  Lines@@@

% L 208, ``\textasciitilde{}'' symbol: This usage of mathematical notation (instead of plain words) in a sentence (here, but also later in the text) is difficult to understand.
%  We removed tilde symbols throughout the manuscript.

% Table 2: Column two (``Methods'') does not provide any information beyond what is explained in ``Description'' and can be removed.
%  We decided to keep the ``methods'' column to make it easier for the reader to refer to the terms used in Mammola et al.~2021 at a glance.

>>>>

\textcolor{blue}{\textbf{Reviewer: 2}}

\textcolor{blue}{Upon reading the title and introduction, I was intrigued and had high expectations for the proposed ``what, how, why'' framework of trait-based analysis.
However, after thoroughly reviewing the manuscript, it became clear that the study does not fully deliver on its initial promise of providing a guideline for applying this framework to ecological data analysis.
Instead, it focuses more on evaluating the varying responses of different functional diversity indices when specific data points are selectively removed from a dataset.
The primary outcome is that different functional diversity indices exhibit distinct and often unpredictable responses to data removal.
However, this observation, while valid, does not seem to offer significant new insights into ecological processes.
In essence, the study highlights the dependence of analytical outcomes on the choice of index - a finding that, while true, feels somewhat self-evident.}

\textcolor{blue}{It is possible that these concerns arise due to insufficient descriptions of the methodology and inconsistencies in terminology (specific points are detailed below).
As the manuscript stands, I am unable to recommend it for publication.
That said, I have outlined several constructive suggestions that I hope will assist the authors in revising and refining the paper.}

\textcolor{blue}{\textbf{General comments}}

\textcolor{blue}{\textbf{Main model idea}}

\textcolor{blue}{The paper falls short in demonstrating how the ``what, how, why'' framework can be applied to ecological data.
This issue may stem from examining two distinct layers: the simple removal of points from datasets and the subsequent interpretation of these removals in terms of ecological processes, which are not explicitly simulated in the model.
The distinction between these layers is confusing and is primarily addressed in Table 1, which does not effectively clarify these points.
To improve clarity, I recommend a more systematic presentation of these issues in the main text: This would involve, describing the relevant ecological processes (e.g., facilitation, filtering, competition), explaining how these processes influence species-associated trait distributions, and then outlining how these influences are represented in the five trait-removal processes.}

@@

\textcolor{blue}{\textbf{Model description:}}
The section on simulating trait-space patterns needs to be significantly expanded to include more technical details regarding how the simulations were conducted.
Key information, such as the model parameters, is entirely absent.
A more thorough explanation of the model setup, parameters, and procedures is required to ensure that the simulations can be reproduced without needing to consult external publications (such as those by Guillerme and Mammola).
To address this, I suggest adding a detailed explanation of the model in either the main Methods section or in a supplementary appendix, to ensure that all necessary parameters and computational steps are clearly specified.}

@@

\textcolor{blue}{\textbf{Consistent terminology:}}

\textcolor{blue}{The paper's clarity is hindered by inconsistent, imprecise, and shifting terminology, making it unnecessarily difficult to follow.
For example, the entities generated by the model are referred to interchangeably as ``elements,'' ``traits,'' ``tips,'' or ``species,'' with no clear rationale for the variation.
Similar, for many other terms that are often used inconsistently or fuzzy (see detailed comments below).
This lack of consistency creates confusion and often forced me to repeatedly revisit sections of the text to discern the intended meaning.
To improve readability, I recommend establishing a consistent terminology early in the paper and adhering to it throughout.
This would significantly reduce ambiguity and enhance the reader's ability to follow the methodology and conclusions.}

% We have now standardised the terminology following \textbackslash cite\{guillerme2020shifting\} and \textbackslash cite\{mammola2021concepts\}
% In brief, elements, tips or species are now referred to as ``observations''; and traits, dimensions and columns as ``dimensions'' when referring to our simulation approach
% We also added a Glossary to the supplementary materials @@@ for clarity.

\textcolor{blue}{\textbf{Detailed comments:}}

\textcolor{blue}{L114: Really ``A single Brownian motion trait''? This implies that only a single trait dimension is generated, whereas from the later text it seems that higher-dimensional traits are simulated.}
%  By ``trait'' here we refer to the single process (Brownian Motion) which can be simulated in any number of dimensions (i.e.~generating a time dependent element value using 2, 4 or 8 dimension coordinates to describe it)
% In this context we differentiate ``dimensions'' which can be interpreted as a column in the data matrix, and the ``trait'' which is the definition of the measured (or simulated) feature
% For example, the trait ``geographical location'' can be described using two dimensions (latitude and longitude) or can be described as two traits (``latitude coordinate'' and ``longitude coordinate'')
% In this manuscript we preferred using the first definition
% We've clarified this as follows:
%  \textit{First, we simulated a single random time dependent trait (Brownian Motion) under a model where lineages only speciate (no extinction; i.e.~a pure birth speciation model) until reaching 200 elements in \textbackslash texttt\{R\} \textbackslash citep\{rcore\} using \textbackslash texttt\{treats\} \textbackslash citep\{guillerme2024treats\}
% This trait was simulated either as a 2, 4 or 8 dimensional trait, in other words a Brownian Motion simulated in a 2, 4 or 8 dimensional space.}
%  Lines@@@

\textcolor{blue}{L115: what is a ``tip''?}
%  By tip we meant a node with no descendants in an evolutionary tree
% We have now standardised the terminology (see response to general comment above).

\textcolor{blue}{L115, ``200 elements'': Why are so few data points used? A dataset of only 200 elements (reduced to 40 after data removal) seems exceptionally sparse, especially in a 4- or 8-dimensional space, as noted later in the text.
What is the motivation behind studying such a sparse space? One of the main advantages of using simple models like a branching process is that they allow for the generation of larger datasets, which would provide more robust statistics and better analytical power.}
%  @@@TODO: Rerun the simulations with 400, 800 and 1000 elements.
%  But also highlight that 40 species is already quite big
% We can also link that to the use of FD in palaeontology (disparity) where 40 is also already quite big!
% Tom's comment: *Reviewer 2's point about why we use only 200 data points - it's a while since I read our paper, but I guess we did that to make it more realistic in terms of the sample sizes people use. In island studies, 200 is pretty large, and I have many references (including my own ...) where people calculate FD using 1-5 species. So again we could make the argument that it represents a trade-off, between realistic sample sizes that people analyse vs large enough that we can really test the mechanisms

\textcolor{blue}{LL 120-124, ``four different stressors, five specific algorithms'': Inconsistency in the used terminology.
In this section, ``stressors'' appear to refer to the amount of data being removed (ranging from 20\% to 80\%), but later, the term ``stressors'' is used to describe the five data-removal processes---referred to as ``specific algorithms'' here.}
% We have now standardised the terminology throughout the manuscript (and included it in the glossary table in the supplementary materials): ``algorithm'' now designate the code algorithm, ``stressors'' (and ``stressors' intensity'') now designates how we modify the data (using the algorithms mentioned above) and ``mechanisms'' now designates the biological mechanisms we are approximating using a specific stressor (that is based on a specific algorithm)
% We've namely changed this section for clarity:
%  \textit{We then applied five different stressors to the trait space with different intensities.}
%  \textit{The five different stressors are effectively five specific algorithms removing either 20\textbackslash\%, 40\textbackslash\%, 60\textbackslash\% or 80\textbackslash\% of the data (the intensities of the stressors, resulting in trait spaces with, respectively, 160, 120, 80 and 40 observations).}
%  \textit{We used the following specific algorithms (Figure \textbackslash ref\{Fig:simulations\}; Table \textbackslash ref\{Tab:mechanisms\}; all algorithms, except ``increasing evenness'\,', were previously described in \textbackslash cite\{guillerme2020shifting\}):}
%  Lines @@@

\textcolor{blue}{L130, ``increasing density'': What is the motivation for this term? While I understand that this removal scheme may result in an overall increase in average density, the removal itself is not directly based on density but rather on pairwise distance in trait space.
For instance, the equalizing scheme would also lead to an increase in density.}
%  We used the term ``increasing density'' here to differentiate from ``decreasing size''
% The reviewer is totally correct in their understanding that, especially for normally distributed data, changes in density in the data actually occur to some extent with all the algorithms used here
% The stressor names and their algorithms are based on Guillerme et al 2020 and their implementation in the dispRity package: in their description they don't assume any specific distribution of the data a priori
% In fact, when using a uniform distribution, the decreasing size stressor does not change the density and the increasing density algorithm does not decrease the size of the trait space.
%  We've now added the following note to include the reviewer's comment:
%  \textit{Note that the algorithm used here to simulate the stressors depend, to some extent, to the distribution of the data for each trait space.}
%  \textit{The algorithms used can lead to similar stressors effects depending on the simulated data distribution.}
%  \textit{For example, ``increasing density'\,' and ``decreasing size'\,' stressors will lead to just what their name suggest if the observations are uniformly distributed (just average increase in density for ``increasing density'\,' and just overall decrease in size for the ``decreasing size'\,' stressors) but will have more or less similar effects (albeit no identical) in the case of a normal distribution (average increase in density and overall decrease in size for both ``increasing density'\,' and ``decreasing size'\,' stressors.)}
%  Lines @@@

\textcolor{blue}{L134, ``shifting space'': Similar to the previous point, this term is somewhat misleading.
This removal scheme does not explicitly shift the distribution; rather, it cuts the distribution at one edge, causing the mean to shift.
However, this effect of shifting the mean is not unique to this scheme - other removal schemes will generally result in a shift of the mean as well.}
%  This is also a fair point
% See comment above.

\textcolor{blue}{L135, ``with the maximal value'': What does ``maximal value'' mean for points in a higher-dimensional space? Maybe this refer to the location in trait space with the highest density of trait elements?}
%  We have now specified what we meant by maximal value:
%  \textit{Shifting space: by removing the required amount of data from a distance (radius) \(
% rho'\) of the observation with the maximum numerical value on all dimensions.}
%  \textit{This is similar to the decreasing size algorithm but instead of choosing \(
% rho\) to be the radius from the centre of the trait space, \(
% rho'\) is a radius from the ``top right corner'\,' of the trait space.}
%  \textit{In other words the centre of the radius \(
% rho'\) is the observation with the highest numerical value on all dimensions - in a 2D representation this is the observation in the top right corner (with \(
% rho'\_{20} \> 
% rho'\_{40} \> 
% rho'\_{60} \> 
% rho'\_{80}\)).}
%  Lines @@@

\textcolor{blue}{LL138- 145: I found this section on the competition method difficult to follow:}

\textcolor{blue}{- The term ``bandwidth'' is unclear.
What does ``bandwidth'' refer to in this context? Does it imply that resampling was done using a frequency-based method, or is it referring to a specific range or interval of trait values in the distribution? This ambiguity ties back to my earlier request for more technical details regarding the simulation process.}

\textcolor{blue}{- I am also unclear on the meaning of ``n''.
Could you clarify what ``n'' represents here? For example, what is meant by the ``probability of sampling of n'' and the reference to ``high/low values of n''? Without this information, it is difficult to grasp how the competition process is functioning.}

\textcolor{blue}{- The term ``flattening the curve'' needs more context.
What curve is being flattened, and how does this relate to the competition method?}

%  We have now completely rewritten the description of this stressor by defining more clearly what the bandwidth is, renaming the variables to be easier to understand, removed mention of flattening the curve and rewrote the equations to be specific to the increase in evenness (previous equations were generalised to both increase or decrease the eveness depending on the variable \textit{i} (formerly \textit{n}) being now always a proportion of removal (e.g.~0.2, 0.4, \ldots{} - former description worked with \textit{n} being also a proportion to add e.g.~1.2, 1.4, etc\ldots):
%  \textit{Increasing evenness: by resampling the proportion of data (i.e.~20\textbackslash\%, 40\textbackslash\%, 60\textbackslash\% or 80\textbackslash\%) but with skewed resampling probabilities.}
%  \textit{In brief, this algorithm reduces the probability of resampling observations in regions of the trait space that have many observations and increases it in regions that have few observations.}
%  \textit{This is done by selecting of \(B\) discrete categories to summarise the distribution - i.e.~bandwidths - using Silverman's ``rule of thumb'\,' \textbackslash citep{[}\textbackslash texttt\{bw.nrd0()\} function in \textbackslash texttt\{R\};{]}{[}{]}\{silverman1986density\}.}
%  \textit{Each of these discrete category \(B\) has an observed probability of sampling of \(b\) (the proportion of observations in the category) and each observation then gets a probability of resampling scaling with the intensity of the stressor (the proportion of what to remove \(i\)) of \(i 
% times (1-b)^{p}\).}
%  \textit{Where \(p\) is factor increasing the scaling power of the algorithm (here we used \(p=3\)).}
%  \textit{In other words, observations in categories with a low density (\(b\<0.5\)) were more likely to be resampled and the ones in categories with high density (\(b\>0.5\)) were less likely to be resampled.}
%  \textit{This approximates our \textbackslash textbf\{competition mechanism\}, were observations in dense regions of the trait space are more likely to be removed than in sparse regions of the trait space.}
%  Lines @@@

\textcolor{blue}{Figure 1: The caption is confusing and needs significant improvement.
Below are some specific points that require attention:}

\textcolor{blue}{- The caption completely omits any description of what is shown in left and right columns (i.e., 1d and 2d case).}
%  We have now clarified what each column contains in the figure caption.
\textcolor{blue}{- It is unclear whether this figure (both the 1d and 2d cases) is based on the actual simulations or is merely a schematic sketch.
For instance, in the right column, it seems unlikely that the figure actually depicts 200 points.}

% The figures does depict 200 points (200 blue in row A and 100 blue + 100 orange in the other columns).
% We have now specified this in the caption.
\textcolor{blue}{- The caption mentions a ``trait space of 2, 4, or 8 dimensions,'' yet a 1d trait space is shown.
If a 1d trait space is an illustrative example, this should be specified in the caption.}
%  We have now specified that we are showing 1 or 2 dimensions regardless of the number of dimensions per traits (i.e.~8 dimensional traits are not fully represented here but what's happening on their first and second dimensions is displayed here)
% We think one confusion on our part is the distinction between a dimension (a column in a matrix) and a trait (a measurable feature for an observation)
% In our case we are always only looking at one trait per trait space
% Although these traits can have 2, 4 or 8 dimensions
% We have now clarified this in the glossary.
\textcolor{blue}{- The pattern shown in the subfigure labeled ``2d facilitating mechanism'' appears questionable.
If point-pairs from a 2d distribution are removed based on a threshold distance, the resulting pattern would not typically resemble what is depicted here.}
%  The illustration of the 2D facilitating mechanism directly comes from the output of the dispRity::reduce.space function where the algorithm is implemented
% Here it selected every \textit{n} pairs of observations that are the closest to each other so that 50\% of the data is selected
% Here \textit{n} is 50 so that 100 observations are selected
% The algorithm is recursive so that once a pair of observations is selected, it cannot be selected again.
\textcolor{blue}{- Terminology: The use of the term ``species'' in the caption is confusing, as the methods seem to generate traits rather than species.}
%  We have now removed any mention of species.
\textcolor{blue}{- The phrase ``removing half of the species, resulting in two groups of tips'' is unclear.
How are these two groups defined? Presumably, one group consists of the removed traits and the other of non-removed traits.
Additionally, it is unclear which color (blue or orange) corresponds to which group.}
%  We have now specified this in the new caption.
\textcolor{blue}{- I am skeptical about the use of ecological terminology such as ``competition'' and ``facilitation'' in this context.
What is actually being shown are data distributions resulting from different point-removal algorithms, not ecological interactions per se.
While these processes can be discussed and associated with ecological mechanisms in the text or discussion (if well justified), using such terms directly in the figure caption is highly misleading.}
%  @@@TODO: Links back to main comment from reviewer 2: Go back to the previous version of table 1 that gives more description of the models
% One thing here is probably to justify more why we think that, for example, ``removing points from the side of a traitspace = filtering''.
%  We now changed the labels in the figure and the caption to highlight that we are indeed approximating such mechanisms, not simulating them per se as the reviewers pointed out.

\textcolor{blue}{LL 150-152: The description is unclear and appears to use the term ``simulation'' inconsistently.
In the phrase ``replicates per simulations,'' the term ``simulation'' seems to have a different meaning than in ``smallest number of simulations.'' Additionally, the first part of the sentence suggests that variance decreases with an increasing number of simulations, whereas the second part implies that adding more simulations leads to increased variance.}
%  There was a confusion indeed
% We now refer to ``the simulation pipeline'' for the whole set of replicates per stressors, and ``replicates'' for each individual run (one trait space + one stressor)
% We changed this sentence to:
%  \textit{We used 171 replicates because that was empirically the smallest number of replicates required to reach a variance between replicates lower than 1\textbackslash\% across all metrics (i.e.~any additional replicates beyond 171 added less that 1\textbackslash\% extra variance).}
%  Lines @@@

\textcolor{blue}{L 166: what does ``uncorrelated'' mean here? Does this mean that the branching process was independently repeated four times, once for each trait dimension, or does it mean that it generates four trait dimensions simultaneously, with these dimensions being uncorrelated?}
%  We meant that the dimensions for each multidimensional traits are uncorrelated (i.e.~the trait is a uncorrelated Brownian motion process in n dimensions)
% We have now clarified this point:
%  \textit{We repeated each simulation pipeline for 4 dimensional traits (generate a trait space and apply the stressor - Figure \textbackslash ref\{Fig:simulation\_results\}) for 2 and 8 dimensional traits with not correlation between each dimensions in each simulation pipeline (supplementary materials).}
%  Lines @@@

\textcolor{blue}{LL 170-178: why then did the authors not simply generate more data points (see my comment above)?}
%  @@@TODO: Rerun the simulations with 400, 800 and 1000 elements.

\textcolor{blue}{Table 1: The integration of this table into the main text could be improved.
Specifically, the abrupt mention of species (which are not part of the model) is confusing.}
%  We have now improved the context of the tables with predictions and removed most of the mentions of ``species'' now replaced by the standardised term of ``observations'' (unless they were specific examples).

\textcolor{blue}{L 178 and several instances later ``o'' should read ``0'' (zero).}
%  These integers were compiled using the latex mathpazo font
% We have now clarified them by changing into a default serif font.

\textcolor{blue}{L184, I suggest to mention here, that these functional metrics are denoted as trait space occupancy later on.}
%  Great suggestions
% We added the following:
%  \textit{We structured our simulations based on three aspects of diversity commonly captured by functional diversity metrics (trait space occupancy metrics; \textbackslash cite\{mammola2021concepts\}) using 12 metrics (Tab
% \textbackslash ref\{Tab:metrics\}):}
%  Lines@@@

\textcolor{blue}{L 208, ``\textasciitilde{}'' symbol: This usage of mathematical notation (instead of plain words) in a sentence (here, but also later in the text) is difficult to understand.}
%  We removed tilde symbols throughout the manuscript.

\textcolor{blue}{Table 2: Column two (``Methods'') does not provide any information beyond what is explained in ``Description'' and can be removed.}
%  We decided to keep the ``methods'' column to make it easier for the reader to refer to the terms used in Mammola et al.~2021 at a glance.
